---
title: "Microbial analysis"
author: "Ares, Angela"
date: "01/06/2022"
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The pipeline followed is this one applied originally for the red soil project in 2018 developed by Maggi Mars: https://maggimars.github.io/RedSoil/16S_Storms_and_RedSoil.html


## Load Packages

```{r packages, echo=FALSE}
library("phyloseq")
library("ggplot2")
library("tidyr")
library("RColorBrewer")
library(reshape2)
library(qiime2R)
library(DESeq2)
library("gridExtra")
library(vegan)
library("metagMisc")
library("grid")
library(jcolors)
library("dplyr")
library("breakaway")
library("CoDaSeq")
library("ggbiplot")
library("intrval")
library("tidyverse")
library("ggpubr")
set.seed(1)
```


Loading the data
```{r}
setwd("~/Documents/GitHub/DNA_monitOki/Amplicon Sequence Analysis")
phyloseq<-qza_to_phyloseq(features="merged_table.qza")
```

Create the metadata file from scratch

```{r}
meta1<- colnames(phyloseq)
metax<- data.frame(meta1)
metadf<- melt(metax)
names(metadf)[1] <- "Sample.ID"
meta2<- cbind(metadf, metadf[1])
names(meta2)[2] <- "SS"
meta3<- meta2 %>%
  separate(SS, c("SS", "Date"), sep = "-", extra = "merge")
meta3<- cbind(meta3, meta3[2])
names(meta3)[4] <- "Area"
meta3<- meta3 %>%
  separate(Date, c("year", "month", "day", "other"), sep = "-") 
meta3$other<- NULL
meta4<- split(meta3,cumsum(1:nrow(meta3)%in%81))
#it creates a list so next we need to create a df out of the list
DF1<- as.data.frame(meta4$`0`)
DF2<- as.data.frame(meta4$`1`)
colnames(DF2) <- c("Sample.ID", "SS", "day", "month", "year", "Area")
meta5<- rbind(DF1, DF2)
meta6<-meta5 %>% unite(date, year, month, day, sep = "-")
meta6$Month<- paste(meta5$month)
#the column Area has some white spaces, we need to get rid of them
meta7<- meta6 %>%
  mutate_if(is.character, str_trim)%>%
  separate(Area, c("Nothing", "LandUse", "Number", "Position"), sep="")
meta7clean <- meta7[, -4] 
meta8<- cbind(meta7clean, meta7clean[4], meta7clean[5])
names(meta8)[8] <- "join1"
names(meta8)[9] <- "join2"
meta8$Area <- paste(meta8$join1, meta8$join2, sep="")
metatable <- meta8[, -c(8,9)] 

#load metadata
row.names(metatable) <- metatable[["Sample.ID"]]
detach("package:dplyr", unload=TRUE)
library("dplyr")
metatable <- metatable %>% select(Sample.ID, everything())
#convert to phyloseq object
Meta_final<- sample_data(metatable)
```

From the qiime2 it is possible to download the .tsv file (16taxonomy) from the viewer and then save it as .csv
```{r}
#loadtaxonomy
taxonomy <- read.csv("16taxonomy.csv", stringsAsFactors = FALSE, header = FALSE)
names(taxonomy) <- c("row", "tax", "Confidence")
row.names(taxonomy) <-taxonomy[[1]]
taxonomy <- taxonomy[,(-1)]
```

```{r}
#SILVA taxonomy is in one column, separate to be able to work with different taxonomic levels:
taxonomy <-  separate(taxonomy, tax, c("D0","D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14"), sep = ";", fill = "right")
#Keep the first 8 taxonomic levels (no assignments afetr that)
taxonomy <- taxonomy[,c(1:8)]
taxmat <- as.matrix(taxonomy)
#covert taxonomy table to phyloseq object
taxon = tax_table(taxmat)

ps = merge_phyloseq(phyloseq, taxon, Meta_final)
```

## Prevalence
Is a measurement which describes in how many samples certain microbes are detected

```{r}
prevdf = apply(X = otu_table(ps),
                MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(ps),
                    tax_table(ps))
#now we plot the prevalence against total abundance

prevplot1<-ggplot(prevdf, aes(TotalAbundance, Prevalence / nsamples(ps),color=D1)) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +  geom_point(size = 2, alpha = 0.7) +
  theme_bw()+
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~D1) + theme(legend.position="none")

prevplot1
```

## Let's do some summary statistics

First we stract ASV table from phyloseq object

```{r}
OTUs <- data.frame(otu_table(ps))
#total number of ASVs in the data set
OTUsRS<- OTUs
OTUsRS$RowSum <- rowSums(OTUsRS)
OTUsRSnoZero <- OTUsRS$RowSum!=0
sum(OTUsRSnoZero)
#Total number of ASVs per sample (range and mean):
OTUs0 <- OTUs!=0 #is this number not a zero? true (1) or false (0)
csums <- colSums(OTUs0) # col sums = observed ASV richness
csumdf <- as.data.frame(csums)
max(csumdf$csums) 
min(csumdf$csums)
mean(csumdf$csums)

#Import denoising stats for both sequencing runs and merge them into one table:
#RSJ1denoise <- read.table("16S_denoisingData.tsv", header = TRUE)
denoise <- read.table("16S_denoisingData.tsv", header = TRUE)
#denoise<- rbind(RSJ1denoise,Junedenoise)

#numbers before denoise
sum(denoise$input)
min(denoise$input)
max(denoise$input)
mean(denoise$input)
#numbers after denoise
sum(denoise$non.chimeric)
min(denoise$non.chimeric)
max(denoise$non.chimeric)
mean(denoise$non.chimeric)
```
## Relative abundance plots
Keep the taxa that are abundant enough to be visible in relative abundance plot:
```{r}
highPrev<-  c("D_1__Actinobacteria", "D_1__Bacteroidetes", "D_1__Chlamydiae;", "D_1__Chloroflexi","D_1__Cloacimonetes", "D_1__Cyanobacteria", "D_1__Dadabacteria","D_1__Dependentiae", "D_1__Epsilonbacteraeota", "D_1__Euryarchaeota", "D_1__Firmicutes", "D_1__Fusobacteria", "D_1__Lentisphaerae", "D_1__Marinimicrobia (SAR406 clade)", "D_1__Patescibacteria", "D_1__Planctomycetes", "D_1__Proteobacteria", "D_1__Verrucomicrobia")
#in total 22 phylla
psNHighPrev<- subset_taxa(ps, D1 %in% highPrev)

#convert counts to relative abundance
physeqPra<- transform_sample_counts(psNHighPrev, function(OTU) 100* OTU/sum(OTU))
```
### Make the plot for field samples

First at Phylum level

```{r}
#D1 taxonomic level
glomD1<- tax_glom(physeqPra, "D1")
#subset to include only samples from the Red Soil project and leave out those collected from pumice stones
psDNA_RS<- subset_samples(glomD1, LandUse == "R" | LandUse == "U")

#Before merging the samples into different groups of aggregation it is essential to transform the variables to factors and then numeric, otherwise the function is not going to work. We can see that the variables are characters

str(sample_data(psDNA_RS))
df <- as.data.frame(lapply(sample_data(psDNA_RS),function (y) if(class(y)!="factor") as.factor(y) else y),stringsAsFactors=T)
row.names(df) <- sample_names(psDNA_RS)
sample_data(psDNA_RS) <- sample_data(df)

#Now most of the variables are factors
str(sample_data(psDNA_RS))

psDNA_RS<-merge_samples(psDNA_RS,"SS")
psDNA_RS<- transform_sample_counts(psDNA_RS, function(OTU) 100* OTU/sum(OTU))

colours= c("#332288", "#44AA99", "#88CCEE", "#882255", "#FFCC00", "#CC6677", "#EE3377", "#7F8E39", "#AA4499", "#BBCCEE", "#222255", "#CCEEFF", "#570D32", "#E48F1B", "#FFCCCC","#D33B44","#1d3554")

#other cool colors("#999933","#DFE07C","#DDAA33","#DDCC77","#CC3311")

#labels for SS (during merging everything was changed)
SSnames<- c("R1_C", "R1_N", "R1_S", "R2_C", "R2_N", "R2_S","U1_C", "U1_N", "U1_S", "U2_C", "U2_N", "U2_S")
#labels for LandUse
LandUselab<- list("1"="Rural", "2"="Urban")
LandUselab_labeller <- function(variable,value){
  return(LandUselab[value])
}
  
generalbarplotD1<-plot_bar(psDNA_RS, x= "SS", fill = "D1",facet_grid= ~LandUse) +  
  scale_y_continuous(expand = c(0, 0)) + 
  ggtitle("") + 
  scale_fill_manual(values=colours ) + 
  scale_x_discrete(labels=SSnames)+ 
  theme(legend.title=element_blank()) + 
  geom_bar(aes(fill=D1), stat="identity", position="stack", width =0.9) +
  theme_classic() + 
  theme(text = element_text(size=14))+
  theme(axis.text.x = element_text(angle = 90)) + 
  facet_grid(~LandUse,scales="free", labeller = LandUselab_labeller) + 
  xlab("Sample") +ylab("Relative Abundance(%)") + 
  theme(text = element_text(size=14))


generalbarplotD1

```

It is clear that some phylla like Epsilonbacteraerota, Firmicutes, Patescibacteria are more present in urban areas than rural ones but it is this tendency the same trough the whole year?
Let's check for different months:
```{r}
#We need to repeat some steps from before so I'll skip the explanations. We just need to change the number when subsetting the month so we can get a plot from each month
psDNA_month<- subset_samples(glomD1, Month == "12")

df <- as.data.frame(lapply(sample_data(psDNA_month),function (y) if(class(y)!="factor") as.factor(y) else y),stringsAsFactors=T)
row.names(df) <- sample_names(psDNA_month)
sample_data(psDNA_month) <- sample_data(df)

str(sample_data(psDNA_month))

psDNA_month<-merge_samples(psDNA_month,"SS")
psDNA_month<- transform_sample_counts(psDNA_month, function(OTU) 100* OTU/sum(OTU))
colours= c("#332288", "#44AA99", "#88CCEE", "#882255", "#FFCC00", "#CC6677", "#EE3377", "#7F8E39", "#AA4499", "#BBCCEE", "#222255", "#CCEEFF", "#570D32", "#E48F1B", "#FFCCCC","#D33B44","#1d3554")

#other cool colors("#999933","#DFE07C","#DDAA33","#DDCC77","#CC3311")

#labels for SS (during merging everything was changed)
SSnames<- c("R1_C", "R1_N", "R1_S", "R2_C", "R2_N", "R2_S","U1_C", "U1_N", "U1_S", "U2_C", "U2_N", "U2_S")

#labels for LandUse
LandUselab<- list("1"="Rural", "2"="Urban")
LandUselab_labeller <- function(variable,value){
  return(LandUselab[value])
}
  
montlybarplotD1<-plot_bar(psDNA_month, x= "SS", fill = "D1",facet_grid= ~LandUse) +  
  scale_y_continuous(expand = c(0, 0)) + 
  ggtitle("") + 
  scale_fill_manual(values=colours ) + 
  scale_x_discrete(labels=SSnames)+ 
  theme(legend.title=element_blank()) + 
  geom_bar(aes(fill=D1), stat="identity", position="stack", width =0.9) +
  theme_classic() + 
  theme(text = element_text(size=14))+
  theme(axis.text.x = element_text(angle = 90)) + 
  facet_grid(~LandUse,scales="free", labeller = LandUselab_labeller) + 
  xlab("Sample") +ylab("Relative Abundance(%)") + 
  theme(text = element_text(size=14))


montlybarplotD1
```

## Alpha Diversity

```{r}
alphazero<-  subset_samples(ps, LandUse == "R" | LandUse == "U")

ad <- breakaway(alphazero)

adiver<- summary(ad) %>%
  add_column("Sample.ID" = alphazero %>% otu_table %>% sample_names)

badf<- merge(adiver, metatable, by = "Sample.ID")

#just comparing sampling sites
SSAD <- ggplot(badf, aes(x=SS, y=estimate, fill = SS)) +  facet_grid(. ~ Month, scales="free" )+ geom_boxplot() + theme_bw() + theme(text = element_text(size=14)) +ylab("Richness Estimate") +xlab("") + scale_fill_manual(values=c("#004488", "#D33B44"))+ theme(text = element_text(size=18)) + theme(legend.position="none")

SSAD

LandUseAD <- ggplot(badf, aes(x=SS, y=estimate, fill = LandUse)) +  facet_grid(. ~ Month, scales="free" )+ geom_boxplot() + theme_bw() + theme(text = element_text(size=14)) +ylab("Richness Estimate") +xlab("") + scale_fill_manual(values=c("#004488", "#D33B44"))+ theme(text = element_text(size=18)) + theme(legend.position="none")

LandUseAD
```

## Ordination and distance

We calculate the Atchison Distance

First, lets perform a log-ratio normalization:

```{r}

OTU4clr<- data.frame(t(data.frame(otu_table(ps))))
row.names(OTU4clr) <- gsub("\\.", "", row.names(OTU4clr))
OTUs.clr <- codaSeq.clr(OTU4clr + 0.5, samples.by.row=TRUE)
OTU2 <- otu_table(as.matrix(OTUs.clr), taxa_are_rows = FALSE)

metatable2<-metatable
row.names(metatable2) <- gsub("-", "", row.names(metatable2))
META2<- sample_data(metatable2)

psCLR <- phyloseq(OTU2,taxon,META2)
```

PCA for all the samples
```{r}
#remove pumice stones samples
psCLR_RS1 <-  subset_samples(psCLR, LandUse != "p" ) 

OTU<- data.frame(otu_table(psCLR_RS1))
metaF <- metatable2[row.names(metatable2) %in% row.names(OTU),]

adonis2(vegdist(OTU , method = "euclidean") ~ SS, data = metaF)
adonis2(vegdist(OTU , method = "euclidean") ~ Month, data = metaF)
```

PCA for all the samples
```{r}
distmat<- vegdist(OTU, method = "euclidean") 

PC <- princomp(distmat)
library("factoextra")
fviz_eig(PC, geom = "bar", bar_width = 0.4) + ggtitle("")
```

Plot with colors indicating month
```{r}
metaF$id<- row.names(metaF)
metaF <- metaF[match(names(PC$scale), metaF$id),]
pMonth<-ggbiplot(PC, var.axes = FALSE, obs.scale = 1, groups = metaF$Month) +theme_bw()+scale_color_manual(labels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"), values=c("#332288", "#44AA99", "#88CCEE", "#882255", "#FFCC00", "#CC6677", "#EE3377", "#7F8E39", "#AA4499", "#BBCCEE", "#222255", "#CCEEFF"), name = "") + geom_point(aes(color=metaF$Month,size = 4))+ theme(text = element_text(size=14)) +ggtitle("") + scale_size_identity()

pMonth
```

Plot with colors indicating Area
```{r}
metaF$id<- row.names(metaF)
metaF <- metaF[match(names(PC$scale), metaF$id),]
parea<-ggbiplot(PC, var.axes = FALSE, obs.scale = 1, groups = metaF$Area) +theme_bw()+scale_color_manual(labels = c("Tancha", "Ogimi", "Ginowan", "Nago"), values=c("#332288", "#44AA99", "#FFCC00", "#CC6677"), name = "") + geom_point(aes(color=metaF$Area,size = 4))+ theme(text = element_text(size=14)) +ggtitle("") + scale_size_identity()

parea
```

Plot with colors indicating SS
```{r}
metaF$id<- row.names(metaF)
metaF <- metaF[match(names(PC$scale), metaF$id),]
pSS<-ggbiplot(PC, var.axes = FALSE, obs.scale = 1, groups = metaF$SS) +theme_bw()+scale_color_manual(values=c("#332288", "#44AA99", "#88CCEE", "#882255","#222255", "#CCEEFF", "#FFCC00", "#CC6677", "#EE3377", "#7F8E39", "#AA4499", "#BBCCEE"), name = "") + geom_point(aes(color=metaF$SS,size = 4))+ theme(text = element_text(size=14)) +ggtitle("") + scale_size_identity()

pSS
```










## Other diversity indexes
```{r}
GP <- prune_species(speciesSums(ps) > 0, ps)

plot_richness(GP, x=SS, measures = c("Chao1", "Shannon"))
```

