---
title: "Microbial analysis"
author: "Ares, Angela"
date: "01/06/2022"
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The pipeline followed is this one applied originally for the red soil project in 2018 developed by Maggi Mars: https://maggimars.github.io/RedSoil/16S_Storms_and_RedSoil.html


## Load Packages

```{r packages, echo=FALSE}
library("phyloseq")
library("ggplot2")
library("tidyr")
library("RColorBrewer")
library(reshape2)
library(qiime2R)
library(DESeq2)
library("gridExtra")
library(vegan)
library("metagMisc")
library("grid")
library(jcolors)
library("dplyr")
library("breakaway")
library("CoDaSeq")
library("ggbiplot")
library("intrval")
library("tidyverse")
library("ggpubr")
set.seed(1)
```


Loading the data
```{r}
setwd("~/Documents/GitHub/DNA_monitOki/Amplicon Sequence Analysis")
phyloseq<-qza_to_phyloseq(features="merged_table.qza")
```

Create the metadata file from scratch

```{r}
meta1<- colnames(phyloseq)
metax<- data.frame(meta1)
metadf<- melt(metax)
names(metadf)[1] <- "Sample.ID"
meta2<- cbind(metadf, metadf[1])
names(meta2)[2] <- "SS"
meta3<- meta2 %>%
  separate(SS, c("SS", "Date"), sep = "-", extra = "merge")
meta3<- cbind(meta3, meta3[2])
names(meta3)[4] <- "Area"
meta3<- meta3 %>%
  separate(Date, c("year", "month", "day", "other"), sep = "-") 
meta3$other<- NULL
meta4<- split(meta3,cumsum(1:nrow(meta3)%in%81))
#it creates a list so next we need to create a df out of the list
DF1<- as.data.frame(meta4$`0`)
DF2<- as.data.frame(meta4$`1`)
colnames(DF2) <- c("Sample.ID", "SS", "day", "month", "year", "Area")
meta5<- rbind(DF1, DF2)
meta6<-meta5 %>% unite(date, year, month, day, sep = "-")
meta6$Month<- paste(meta5$month)
#the column Area has some white spaces, we need to get rid of them
meta7<- meta6 %>%
  mutate_if(is.character, str_trim)%>%
  separate(Area, c("Nothing", "LandUse", "Number", "Position"), sep="")
meta7clean <- meta7[, -4] 
meta8<- cbind(meta7clean, meta7clean[4], meta7clean[5])
names(meta8)[8] <- "join1"
names(meta8)[9] <- "join2"
meta8$Area <- paste(meta8$join1, meta8$join2, sep="")
metatable <- meta8[, -c(8,9)] 

#load metadata
row.names(metatable) <- metatable[["Sample.ID"]]
detach("package:dplyr", unload=TRUE)
library("dplyr")
metatable <- metatable %>% select(Sample.ID, everything())
#convert to phyloseq object
Meta_final<- sample_data(metatable)
```

From the qiime2 it is possible to download the .tsv file (16taxonomy) from the viewer and then save it as .csv
```{r}
#loadtaxonomy
taxonomy <- read.csv("16taxonomy.csv", stringsAsFactors = FALSE, header = FALSE)
names(taxonomy) <- c("row", "tax", "Confidence")
row.names(taxonomy) <-taxonomy[[1]]
taxonomy <- taxonomy[,(-1)]
```

```{r}
#SILVA taxonomy is in one column, separate to be able to work with different taxonomic levels:
taxonomy <-  separate(taxonomy, tax, c("D0","D1", "D2", "D3", "D4", "D5", "D6", "D7", "D8", "D9", "D10", "D11", "D12", "D13", "D14"), sep = ";", fill = "right")
#Keep the first 8 taxonomic levels (no assignments afetr that)
taxonomy <- taxonomy[,c(1:8)]
taxmat <- as.matrix(taxonomy)
#covert taxonomy table to phyloseq object
taxon = tax_table(taxmat)

ps = merge_phyloseq(phyloseq, taxon, Meta_final)
```

## Prevalence
```{r}
prevdf = apply(X = otu_table(ps),
                MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(ps),
                    tax_table(ps))
#now we plot the prevalence against total abundance

prevplot1<-ggplot(prevdf, aes(TotalAbundance, Prevalence / nsamples(ps),color=D1)) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +  geom_point(size = 2, alpha = 0.7) +
  theme_bw()+
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~D1) + theme(legend.position="none")

prevplot1
```

## Let's do some summary statistics

First we stract ASV table from phyloseq object

```{r}
OTUs <- data.frame(otu_table(ps))
#total number of ASVs in the data set
OTUsRS<- OTUs
OTUsRS$RowSum <- rowSums(OTUsRS)
OTUsRSnoZero <- OTUsRS$RowSum!=0
sum(OTUsRSnoZero)
#Total number of ASVs per sample (range and mean):
OTUs0 <- OTUs!=0 #is this number not a zero? true (1) or false (0)
csums <- colSums(OTUs0) # col sums = observed ASV richness
csumdf <- as.data.frame(csums)
max(csumdf$csums) 
min(csumdf$csums)
mean(csumdf$csums)

#Import denoising stats for both sequencing runs and merge them into one table:
#RSJ1denoise <- read.table("16S_denoisingData.tsv", header = TRUE)
denoise <- read.table("16S_denoisingData.tsv", header = TRUE)
#denoise<- rbind(RSJ1denoise,Junedenoise)

#numbers before denoise
sum(denoise$input)
min(denoise$input)
max(denoise$input)
mean(denoise$input)
#numbers after denoise
sum(denoise$non.chimeric)
min(denoise$non.chimeric)
max(denoise$non.chimeric)
mean(denoise$non.chimeric)
```
## Relative abundance plots
Keep the taxa that are abundant enough to be visible in relative abundance plot:
```{r}
highPrev<-  c("D_1__Actinobacteria", "D_1__Bacteroidetes", "D_1__Chlamydiae;", "D_1__Chloroflexi","D_1__Cloacimonetes", "D_1__Cyanobacteria", "D_1__Dadabacteria","D_1__Deinococcus-Thermus", "D_1__Dependentiae", "D_1__Epsilonbacteraeota", "D_1__Euryarchaeota", "D_1__Firmicutes", "D_1__Fusobacteria", "D_1__Kiritimatiellaeota", "D_1__Lentisphaerae", "D_1__Marinimicrobia (SAR406 clade)", "D_1__Patescibacteria", "D_1__Planctomycetes", "D_1__Proteobacteria","D_1__Synergistetes", "D_1__Thaumarchaeota", "D_1__Verrucomicrobia")
#in total 22 phylum
psNHighPrev<- subset_taxa(ps, D1 %in% highPrev)

#convert counts to relative abundance
physeqPra<- transform_sample_counts(psNHighPrev, function(OTU) 100* OTU/sum(OTU))
```
### Make the plot for field samples

First at Phylum level

```{r}
#D1 taxonomic level
glomD1<- tax_glom(physeqPra, "D1")
#subset to include only samples from the Red Soil project and leave out those collected from pumice stones
psDNA_RS<- subset_samples(glomD1, LandUse == "R" | LandUse == "U")
#make relative abundance

newcolors= c("#DDCC77","#88CCEE", "#44AA99", "#FFCC00", "#CC3311", "#CC6677", "#FFCCCC", "#999933", "#DDCC77","#EE3377", "#882255", "#AA4499", "#BBCCEE", "#222255", "#CCEEFF", "#DDAA33","#332288", "#88CCEE", "#44AA99", "#FFCC00", "#CC3311", "#CC6677")

metaDNA_RS <- metatable[metatable$LandUse =="R" | metatable$LandUse == "U", ]
#metaFieldRS$Treatment <- factor(metaFieldRS$Treatment, levels=c("A1 6/13", "A2 6/13", "A3 6/13", "A4 6/13", "A1 6/16", "A2 6/16", "A3 6/16", "A4 6/16", "A1 6/19", "A2 6/19", "A3 6/19", "A4 6/19", "RS1", "RS2", "RS3", "RS4", "A1 9/28", "A2 9/28", "A3 9/28", "A4 9/28", "A1 10/01", "A2 10/01", "A3 10/01", "A4 10/01", "A1 10/03", "A2 10/03", "A3 10/03", "A4 10/03", "A1 10/08", "A2 10/08", "A3 10/08", "A4 10/08", "RS 2", "RS 3" ))
METArs<- sample_data(metaDNA_RS)
sample_data(psDNA_RS) <- METArs


taxabarplotD1<-plot_bar(psDNA_RS, x= "SS", fill = "D1", facet_grid= ~Month) +  
  scale_y_continuous(expand = c(0, 0)) + 
  ggtitle("") + 
  scale_fill_manual(values=newcolors ) + 
  theme(legend.title=element_blank()) + 
  geom_bar(stat="identity", position="stack", width =0.9) +
  theme_classic() + 
  theme(text = element_text(size=14))+
  theme(axis.text.x = element_text(angle = 90)) + 
  xlab("Sample") +ylab("Relative Abundance(%)") + 
  facet_grid(~Month,scales="free") + 
  theme(text = element_text(size=14))

taxabarplotD1+ theme(legend.position="none")
```


